Data Appendix to \"The impact of weather events on schooling outcomes\

### 2020-03-19 data_appendix.Rmd
updated text

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")
# input Stata file
library(foreign)
data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")

#The following custom function simplifies the process of writing dfSummaries to html files
export_summary_table <- function(dfSummary_output){
  data_info <- attr(dfSummary_output, "data_info")
  ds_name <- data_info$Data.frame
  print(dfSummary_output,
      file = str_c("output/", ds_name, "_summary.html"),
      method = "browser",
      report.title = ds_name)
}
```

@@

*Note: this document structure will require you to re-run steps of your analysis multiple times. If your code takes a long time, please come talk with me about strategies to reduce run time or save earlier results.*

# Instructions for Use
This document includes instructions for how to create your Data Appendix. Outside of this section, instruction paragraphs are listed in *italics* (like the first paragraph above). Instructions should be removed before submission.

@@

# Raw data
*Each dataset you use will have its own documentation section. The next subsection in this document (Dataset description) is a template. You can copy this section and paste it into your document each time you need to add a section for a new dataset. Note that each line in the Dataset description section __must__ end with two spaces.* 
This section documents the datasets used in this analysis.

@@

- **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **variable_name2:** Description of second variable.

@@

*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing. After importing, export a dataframe summary using the command.*

`export_summary_table(dfSummary(dataset_name))`

*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output. I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process. Since these files will exist only electronically, their length is less of a concern. If you like to print out files to proofread and want me to help you shorten the printed versions, let me know. We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*

>>>>

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

#The following custom function simplifies the process of writing dfSummaries to html files

export_summary_table <- function(dfSummary_output){
  data_households <- attr(dfSummary_output, "data_info")
  total_2005 <- data_households$hh_2005
  print(dfSummary_output,
      file = str_c("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\", total_2005, "_summary.html"),
      method = "browser",
      Weather.and.school = total_2005)
}
```

# Raw data
This section documents the datasets used in this analysis: household data, flood data, and drought data from 2005 in Bolivia.

```{r input households data, echo = F, message = F}
# input Stata file, add haven package to make stata 16 version readable 
library(foreign)
library(haven)
data_2005 <- read_dta("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\seminar_paper_data\\2005_hhsurvey.dta")
#install.packages("reshape")
require(reshape)
data_2005 <- rename(data_2005, c(folio = "hh_id"))
data_2005 <- rename(data_2005, c(s3_08 = "assist"))
data_2005 <- rename(data_2005, c(s3_09 = "reason"))
data_2005 <- rename(data_2005, c(nm_y = "members"))
data_2005 <- rename(data_2005, c(s1_02 = "sex"))
data_2005 <- rename(data_2005, c(s1_03 = "age"))
data_2005 <- rename(data_2005, c(s1_05 = "relation"))
data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
data_2005 <- rename(data_2005, c(s3_04 = "register"))
data_2005 <- rename(data_2005, c(s3_05b = "grade"))
data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
data_2005 <- rename(data_2005, c(depto = "dept"))
```

```{r floods data set imported, echo = F, message = F}
#floods data set

floods <- read_csv("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\recurrencia_inundacion2002_2012bol.csv")

#rename variables
floods <- rename(floods, c(NOM_DEP = "dept"))
floods <- rename(floods, c(A_2005 = "flood"))

library(dplyr)
floods %>% 
  group_by(dept) %>% 
  summarise(flood = sum(flood))

```

```{r importing droughts data, echo = F, message = F}
#drought data set 
drought <- read_csv("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\recurrenciasequia_2002_2012bol.csv")

#rename variables
drought <- rename(weather, c(NOM_DEP = "dept"))
drought <- rename(weather, c(A_2005 = "drought"))

library(dplyr)
drought %>% 
  group_by(dept) %>% 
  summarise(drought = sum(drought))

```

@@

# Instructions for Use

@@

@@

- **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **nro1:** Description of second variable.
- **assist:** 
- **urb_rur:** 
- **reason:** 
- **members:** 
- **sex:** 
- **age:** 
- **relation:** 
- **literacy:** 
- **register:**
- **grade:** 
- **repeated:** 
- **dept:**

@@

*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing. After importing, export a dataframe summary using the command.*
```{r selecting variables}
# select variables 
hh_vars <- c("hh_id", "nro1", "urb_rur", "assist", "reason", "members", "sex", "age", "relation", "literacy", "register", "grade", "repeated", "dept")
hh_2005 <- data_2005[hh_vars]

data_2005 <- rename(data_2005, c(folio = "hh_id"))
data_2005 <- rename(data_2005, c(s3_08 = "assist"))
data_2005 <- rename(data_2005, c(s3_09 = "reason"))
data_2005 <- rename(data_2005, c(nm_y = "members"))
data_2005 <- rename(data_2005, c(s1_02 = "sex"))
data_2005 <- rename(data_2005, c(s1_03 = "age"))
data_2005 <- rename(data_2005, c(s1_05 = "relation"))
data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
data_2005 <- rename(data_2005, c(s3_04 = "register"))
data_2005 <- rename(data_2005, c(s3_05b = "grade"))
data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
data_2005 <- rename(data_2005, c(depto = "dept"))

#select weather variables
flood_vars <- c("dept", "flood")
flood_2005 <- floods[flood_vars]

```

```{r merging datasets}
total_2005 <- merge(hh_2005, weather_2005,by="dept")
```

```{r exporting data table}
export_summary_table(dfSummary(hh_2005))
```
